---
title: "Team.Rmd"
author: "Jacob Shoemaker; jshoemaker2249, Qinwen Yang; AngelQinwen"
date: "2/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Data Cleaning
```{r}
HAPPY <- readRDS("HAPPY.rds")
```

```{r}
HAPPY[HAPPY == "IAP" | HAPPY == "DK"| HAPPY == "NA"] <- NA
```

```{r}
HAPPY$HAPPY <- as.factor(HAPPY$HAPPY)
HAPPY$SEX <- as.factor(HAPPY$SEX)
HAPPY$MARITAL <- as.factor(HAPPY$MARITAL)
HAPPY$DEGREE <- as.factor(HAPPY$DEGREE)
HAPPY$FINRELA <- as.factor(HAPPY$FINRELA)
HAPPY$HEALTH <- as.factor(HAPPY$HEALTH)
HAPPY$PARTYID <- as.factor(HAPPY$PARTYID)
HAPPY$POLVIEWS <- as.factor(HAPPY$POLVIEWS)
HAPPY$AGE <- as.numeric(HAPPY$AGE)
HAPPY$AGE <- ifelse(HAPPY$AGE == "89 OR OLDER", 89, HAPPY$AGE)
```

```{r}
#Order Degree
HAPPY$DEGREE <- factor(HAPPY$DEGREE, levels = c("LT HIGH SCHOOL", "HIGH SCHOOL", "JUNIOR COLLEGE", "BACHELOR", "GRADUATE"))

#Order Opinion of Financial Standing
HAPPY$FINRELA <- factor(HAPPY$FINRELA, levels = c("FAR BELOW AVERAGE", "BELOW AVERAGE", "AVERAGE", "ABOVE AVERAGE", "FAR ABOVE AVERAGE"))

#Order Health
HAPPY$HEALTH <- factor(HAPPY$HEALTH, levels = c("POOR", "FAIR", "GOOD", "EXCELLENT"))

#Order Political Party
HAPPY$PARTYID <- factor(HAPPY$PARTYID, levels= unique(c("STRONG DEMOCRAT", "NOT STR DEMOCRAT", "IND,NEAR DEM","INDEPENDENT", "IND,NEAR REP", "NOT STR REPUBLICAN", "STRONG REPUBLICAN", "INDEPENDENT")))

#Order Political Views
HAPPY$POLVIEWS <- factor(HAPPY$POLVIEWS, levels = c("EXTREMELY LIBERAL", "LIBERAL", "SLIGHTLY LIBERAL", "MODERATE", "SLIGHTLY CONSERVATIVE", "CONSERVATIVE", "EXTRMLY CONSERVATIVE"))

#Order Happiness Level
HAPPY$HAPPY <- factor(HAPPY$HAPPY, levels = c("NOT TOO HAPPY", "PRETTY HAPPY", "VERY HAPPY"))


#Find average age of each marital status
mean(HAPPY$AGE[HAPPY$MARITAL == "NEVER MARRIED"], na.rm = TRUE)
mean(HAPPY$AGE[HAPPY$MARITAL == "MARRIED"], na.rm = TRUE)
mean(HAPPY$AGE[HAPPY$MARITAL == "SEPARATED"], na.rm = TRUE)
mean(HAPPY$AGE[HAPPY$MARITAL == "DIVORCED"], na.rm = TRUE)
mean(HAPPY$AGE[HAPPY$MARITAL == "WIDOWED"], na.rm = TRUE)

#Order Marital Status in ascending order of these averages
HAPPY$MARITAL <- factor(HAPPY$MARITAL, levels = c("NEVER MARRIED", "SEPARATED", "MARRIED", "DIVORCED", "WIDOWED"))
```

2. Data Exploring

#Investigate the distributions of 2 different variables
```{r}

library(ggplot2)

ggplot(HAPPY, aes(x=HAPPY, fill = SEX))+geom_bar(na.rm=TRUE, color="purple",size=0.5)+ylab('Total Number')+xlab('Happiness')+ggtitle("Relationship between HAPPY and SEX")

#distribution of degree by sex
ggplot(HAPPY, aes(x=HAPPY, fill=FINRELA))+ geom_bar(na.rm = TRUE, color= "black", size = 0.4) + ggtitle("Distribution of Happiness and Opinion of Family Income") + ylab("Count") + xlab("Happiness") + guides(fill=guide_legend(title="Opinion of Family Income"))

```


As you can see here, there is certainly a noticeable difference in happiness depending one someone's sex. More females would consider themselves to be "pretty happy" than males, however this could potentially be explained by the fact that there are more females than males within the dataset. As for the relationship between financial stability and happiness, it would appear that those who are below or far below average in their opinion of their financial success are more likely to be unhappy.


#Investigate the relationship between happiness and three other variables in the data.

```{r}
ggplot(HAPPY, aes(x=HAPPY, fill= DEGREE))+geom_bar(na.rm = TRUE, color="black", size = 0.3)+xlab('Happiness')+ylab('Total Number')+ggtitle("Relationship between Happy and degree")

ggplot(HAPPY, aes(x=HAPPY, y=AGE, fill=SEX)) + geom_boxplot(na.rm = TRUE, size=0.3)+ xlab("Happiness")+ ggtitle("Happiness and Age") + ylab("Age")+coord_flip() 

library(ggmosaic)
ggplot(HAPPY)+geom_mosaic(aes(x=product(HEALTH), fill=HAPPY)) + xlab("Health") + ylab("Happiness") + ggtitle("Mosaic Plot of Health and Happiness") + guides(fill=guide_legend(title="Happiness Level"))
```


Starting with degree, there ratio of people that are unhappy with higher levels of education is certainly lower than those with solely a high school diploma. A potential reason for this is the job opportunities and income associated with more education. From an age standpoint, one could make the case that as a woman gets older, she tends to become more unhappy. Lastly, those in poor health condition tend to gravitate towards being unhappy, which certainly makes sense due to quality of life.


#Final Team Analysis of Happiness
From the dataset, we can't conclude that one factor have crucial effect than any other factors when determing people's happiness. But we are able to conclude that most female are happier than male. Health would effect happiness which poor health people are tend to become less happier, and people who are in good and excellent health are happier than the people who are in poor health. In addition, People who have married is happier than separated and divorced. In terms of financial standing, being below or far below average is more closely connected to being unable to make the jump from pretty happy to very happy. From health, being in poor health conditions is more closely related to being unhappy.


